Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-06 chapters/paper2.tex
Changes 1. Completed rewriting introduction of aim 2

% Stimuli-responsive emulsions have attracted significant interest for applications in enhanced oil recovery, porous material templating, 
% in-situ reaction separation systems, and drug delivery scaffolds \cite{tang_stimuli-responsive_2016, bago_rodriguez_capsules_2019}. 
% These applications are facilitated by stimuli-induced microstructural modifications, which alter the permeability, surface area, and 
% morphology of the emulsion. Advancements in particle synthesis techniques, along with growing concerns over surfactant toxicity and 
% environmental impact, have contributed to the renewed interest in particle-stabilized emulsions since their initial discovery by Ramsden 
% and Pickering in 1907 \cite{rozynek_opening_2019, zhou_magnetic_2011}. Stimuli-responsive behavior in these emulsions has been achieved 
% through changes in pH, temperature, or chemical composition 
% \cite{maingret_dextran-based_2020, lin_ph-responsive_2018, raju_ph-responsive_2018, zhang_ph_2022}.

Due to their tortuous, co-continuous domain morphology, bijels have been 
extensively studied for applications in membrane fabrication, catalyst supports, battery electrodes, and pharmaceutical 
products \cite{yabuno_preparation_2020, samdani_bicontinuous_2017, cha_bicontinuous_2019, garcia_scalable_2019, santiago_cordoba_aerobijels_2020}.  
Bijels form through the arrested spinodal decomposition of partially miscible fluids 
\cite{stratford_colloidal_2005, herzig_bicontinuous_2007, tavacoli_novel_2011}. During phase separation, the emerging interface propagates 
through the system, driving particle adsorption. Once the available interfacial area matches the total cross-sectional area of the adsorbed 
particles, the structure becomes jammed in place \cite{stratford_colloidal_2005, herzig_bicontinuous_2007, tavacoli_novel_2011}. Since their 
initial discovery in simulations and subsequent experimental synthesis, a diverse body of literature has emerged, exploring fluid systems 
compatible with bijel formation, scalable fabrication techniques, and potential applications.

Bijels stabilized by ellipsoidal particles exhibit distinct domain evolution timescales compared to those stabilized by 
spherical particles \cite{gunther_timescales_2014}. Unlike spherical particles, ellipsoidal particle reorientation at the 
interface influences microstructure. After jamming, additional timescales emerge due to capillary interactions that drive 
particle reorientation and prolonged domain coarsening \cite{gunther_timescales_2014}. Bijels behave as 2D glasses percolating 
in 3D space, with locally jammed particles as observed by Günther et al. 
\cite{ching_bijel_2022, torquato_jammed_2010, gunther_timescales_2014}. Studies on jammed suspensions indicate subdiffusive 
particle movement, further confirming local jamming in bijel monolayers \cite{savelev_diffusion_2006}.

Magnetic fields have gained interest as external stimuli due to their targeted interactions, biocompatibility, and material 
versatility. The structural response of magnetically responsive hydroxyapatite crystals depends on initial particle orientation
\cite{nakayama_stimuli-responsive_2018}. Magnetophoresis, where particles move along field lines, has been observed in Pickering 
emulsions stabilized by magnetic microspheres \cite{tham_magnetophoresis_2021}. Structural deformation and de-emulsification occur due 
to particle migration or tilting at the interface \cite{yang_rapid_2020, misra_magnetic_2020}. However, how these effects influence 
bijel microstructure post-jamming remains unclear. Given the locally jammed nature of bijels, magnetic field responses, combined with 
anisotropic particle stabilization, may offer a means to control bijel microstructure.

This work investigates the structural response of bijels stabilized by anisotropic magnetic particles. Using a Lattice Boltzmann Method 
coupled with Molecular Dynamics, we simulate a binary fluid system where ellipsoidal particles, modeled as magnetic dipoles along their 
symmetry axis, stabilize bijel templates. Magnetic fields applied to bijels made without a magnetic field reveal anisotropic bijel formation, 
with microstructure evolution dependent on particle reordering at the interface, quantified using the six-fold 
Steinhardt order parameter $\langle Q6 \rangle$.

We further assess how initial microstructure influences structural changes, showing that the difference in order between initial and final 
configurations of the particle monolayers dictates the extent of microstructure evolution if the applied field is increased in the direction of 
the existing order. Additionally, applying and removing a 
magnetic field results in distinct microstructural properties and timescales, highlighting the importance of processing history. These findings 
demonstrate how magnetic fields, combined with anisotropic particle-stabilized bijels, enable on-demand microstructure switching for potential applications.

>>>>

Emulsions can adopt a variety of microstructures depending on fluid volume fractions and the affinity of stabilizing agents for each phase. Among these, bicontinuous 
interfacially jammed emulsion gels (bijels) are characterized by a tortuous, co-continuous network of immiscible fluid domains stabilized by colloidal particles. This 
unique architecture lends itself to numerous advanced applications, including membrane fabrication, catalyst supports, battery electrodes, and pharmaceutical delivery 
systems. Bijels form through the arrest of spinodal decomposition in partially miscible fluid mixtures: as phase separation proceeds, the interface sweeps through the 
system, adsorbing particles until the interfacial area becomes saturated, resulting in jamming. Microstructural features such as domain size, continuity, and interfacial 
curvature critically influence material performance, affecting properties like mass transport, mechanical stability, and surface reactivity. Since their discovery via 
simulation and subsequent experimental realization, research on bijels has expanded significantly, exploring diverse synthesis conditions and scalable processing methods.

The functional properties of bijels are inherently linked to their internal structure. For instance, lower tortuosities in the direction of ion transport yield improved 
charge rates and cycling stability in battery electrodes \cite{ebner_tortuosity_2014, samdani_bicontinuous_2017}. In liquid-liquid extraction, the characteristic pore 
lengthscale determines molecular selectivity \cite{khan_nanostructured_2022}, while in pharmaceutical applications, domain size influences drug release rates and cellular 
behavior \cite{vanoli_bijels_2022, thorson_bijel-templated_2019}. Given the sensitivity of these properties to microstructural characteristics, the ability to dynamically 
modify the morphology of bijels presents a promising strategy to adapt their behavior in situ for targeted performance.

Stimuli-responsive emulsions have garnered increasing interest for applications requiring dynamic control over material properties, including enhanced oil recovery, reaction 
separation, and therapeutic delivery \cite{tham_magnetophoresis_2021, cui_stabilizing_2013, rozynek_opening_2019, lu_controllable_2020}. These systems operate by using 
external stimuli to alter interfacial configurations, enabling modulation of permeability, interfacial area, and morphology. For example, in emulsions stabilized by 
spherical particles, electric fields can induce droplet deformation via particle unjamming and migration before re-establishing interfacial jamming \cite{cui_stabilizing_2013}. 
Among various external stimuli, magnetic fields are particularly attractive due to their remote addressability, material compatibility, and bio-relevance. 
Core-shell ferrite-hydroxyapatite nanorods, for instance, exhibit reversible morphological changes depending on initial particle orientation and field direction 
\cite{nakayama_stimuli-responsive_2018}. In magnetically stabilized Pickering emulsions, magnetophoresis has been shown to drive particle migration and interfacial 
distortion or collapse \cite{tham_magnetophoresis_2021, yang_rapid_2020, misra_magnetic_2020}. Notably, Karthikeyan and Schiller demonstrated that magnetic fields applied 
during bijel formation can bias domain coarsening, producing anisotropic microstructures via orientation-specific jamming \cite{karthikeyan_formation_2024}. However, the 
structural response of bijels to magnetic fields after formation remains poorly understood.

This study investigates the post-formation structural response of bijels stabilized by ellipsoidal magnetic particles under the influence of externally applied magnetic fields. 
Using a hybrid Lattice Boltzmann Method coupled with Molecular Dynamics, we simulate the formation and evolution of bijel structures stabilized by rod-like particles modeled as 
magnetic dipoles. We find that increasing the applied magnetic field after formation induces significant microstructural changes, including increased domain anisotropy and 
coarsening, up to a saturation threshold. In contrast, decreasing or removing the magnetic field results in minimal morphological change, suggesting a strong hysteresis effect. 
Analysis reveals that particle unjamming, reorientation at the interface, and subsequent re-jamming are responsible for these structural transformations.

To understand the saturation behavior, we applied magnetic fields of varying strength to bijel templates formed under increasing initial field conditions. We observed that 
a greater initial nematic order parameter negatively correlated with the extent of microstructure change. Finally, switching off the magnetic field in pre-aligned bijels had little 
impact on the microstructure, indicating that the jammed interfacial network prevents reversal of anisotropic features. These findings underscore the importance of processing 
history and demonstrate how magnetic field manipulation, in combination with anisotropic particle stabilization, enables tunable, on-demand microstructure control in bijels, 
offering new capabilities for adaptive material design.
